---
title: CFB Worst Games
author: Caleb Badura
date: '2021-10-24'
slug: text-cleaning
categories:
  - R
tags:
  - R Markdown
---

Installed and loaded packages

```{r}
install.packages("ggalt")
```

```{r}
library(tidyverse)
library(ggalt)
```

I am using football data from Sports Reference

```{r}
badlogs <- read_csv("~/Desktop/SPMC350/badfootballlogs19.csv")
```

The first step is to separate the Result column into two columns, which are: "Outcome" and "Score." I also removed the () from the score.  

Since we need to separate the scores, I created two different columns of Team and Opponent Scores by separating the Score column.  

I saved this to the "badlogs19" dataframe. 

```{r}
badlogs %>% 
  separate(Result, into=c("Outcome","Score"), sep="\\(") %>%
  mutate(Score=gsub(")","",Score)) %>%
  separate(Score, into=c("TeamScore","OpponentScore"), sep="-") -> badlogs19
```

I want to find the Differential, so to do that, I mutated a new field by subtracting OpponentScore from TeamScore.  I saved that to a new dataframe called "Diff"

```{r}
badlogs19 %>%  mutate(OpponentScore=as.numeric(OpponentScore)) %>% 
                mutate(TeamScore = as.numeric(TeamScore)) %>%
            mutate(Differential= TeamScore- OpponentScore) %>%
  arrange(desc(Differential)) -> Diff
```

I want to see specific games where the differential was above 65, so I created a new dataframe called "worstgames."

```{r}
worstgames <- Diff %>% filter(Differential > 65)
```

Created a Scatterplot from the data. 

```{r}
ggplot() + geom_point(data=Diff, aes(x=TeamScore, y=OpponentScore), color='grey') +
          geom_point(data=worstgames, aes(TeamScore, y=OpponentScore, color= 'red')) +
        geom_encircle(data=worstgames, aes(x=TeamScore, y=OpponentScore), s_shape=.5, expand=.03, colour="black")+
      labs(x="Opponent Score", y= "Team Score", title= "What Were The Worst Games in College Football?", subtitle= "The worst games in point differential are quite easy to spot.", caption= "Source= Sports-Reference.com | By Caleb Badura") +
  theme_minimal() +
  theme(
    plot.title = element_text(size=15, face="bold")
  )
  
```

